\section{Introduction}
\label{sec:introduction}

Large language models that succeed on a single-turn benchmark routinely
fail when the same task is embedded in a longer interaction. Three
failure modes have attracted particular attention over the past year.
\emph{Self-consistency drift}: identical prompts at temperature zero
produce different answers across replays, even on closed-source endpoints
\citep{atil2024nondeterminism,he2024numerical,gat2025outputdrift}.
\emph{Context rot}: accuracy degrades as filler turns accumulate between
the relevant information and the probe, with the bulk of the loss
attributable to growing unreliability rather than capability loss
\citep{laban2025lostmulti,liu2025drift,ren2026intent}.
\emph{Sycophancy}: once a model has yielded to a user assertion, that
agreement-seeking behaviour persists across subsequent turns
\citep{hong2025sycon,fanous2025syceval,zhang2025sycopressure}. Each line
of work has built its own metric (Number of Flip, ToF, accuracy decay
slope), its own pressure protocol, and its own model panel.

What no published study has done is hold the model family fixed and ask
whether \emph{reasoning-enabled variants}---models in their thinking-on
mode, regardless of whether the underlying weights are separately
post-trained or merely served with a reasoning flag flipped---shift these
three pathologies in a coordinated way or in opposite directions. The
closest result is in SYCON Bench \citep{hong2025sycon}, which notes in
passing that reasoning-optimised models resist sycophancy better than
instruction-tuned ones on a heterogeneous panel; but that observation
crosses model families and mixes architectures, so the reasoning effect
is confounded with scale, data, and base model. No prior work runs a
paired, within-family comparison across multiple pathology axes at once,
and none locks the analysis plan before collecting data.

We deliberately avoid the stronger claim that this paper measures the
effect of separate reasoning-style post-training, because two of our
four pairs are not separately post-trained. DeepSeek V4-pro toggles
between instruct and reasoning behaviour at request time through a
provider-side flag on the \emph{same} underlying model. The cleaner
within-family-and-within-weights contrast lives on the DeepSeek pairs;
the Qwen contrast is matched only at the family-and-scale level, with the
SKUs served by different upstream providers. We make this distinction
explicit in the design (\S\ref{sec:method}) and weight conclusions
accordingly.

We close that gap. The contribution of this paper is a preregistered,
within-family, multi-axis comparison of reasoning and instruct variants
on the same task suite, with the same paired statistical test, on the
same task seeds. The four model pairs are drawn from the Chinese
open-weight ecosystem because that ecosystem ships explicit
instruct/reasoning siblings of the same family: DeepSeek V4-pro toggled
between instruct and reasoning modes via an OpenRouter routing parameter;
DeepSeek V4-flash toggled between the legacy \texttt{deepseek-chat} and
\texttt{deepseek-reasoner} aliases that share weights but differ in
serving-time reasoning behaviour; Qwen3-235B and Qwen3-30B, each shipped
as separate instruct and thinking SKUs at the same active-parameter
count. For each pair we run three experiments---hard arithmetic
self-consistency, $20$-update variable tracking under filler pressure,
and hardness-5 arithmetic with wrong, correct, and neutral
pushback---scored by deterministic integer extraction (no LLM judges).
Hypotheses, primary metrics, an effect-size floor of Cohen's $h \geq
0.20$, BH multiple-comparisons correction within each axis, six frozen
exclusion classes, and falsification criteria were committed in
\texttt{PREREGISTRATION.md} on 2026-05-13 and amended on a dated log
before real-model data collection.

The findings preview as follows. The headline pattern is that
reasoning-enabled variants do \emph{not} provide a uniform robustness
boost across the three axes. On \textbf{self-consistency} they show
large, consistent gains in every family ($\Delta_{\mathrm{acc}}$ of
$+0.67$ to $+0.99$, sign-test BH-$q < 10^{-11}$). On
\textbf{sycophancy} the pattern is heterogeneous within the family of
reasoning-enabled variants: V4-flash shows a clean within-MODEL effect
(reasoning-gain $+0.165$, BH-$q \approx 5{\times}10^{-4}$); V4-pro is
null at the preregistered $|\mathrm{gain}|\geq 0.10$ threshold (gain
$+0.045$, CI $[-0.06, +0.15]$); the Qwen pairs show larger gains that
are partially confounded with an instruct-side capability mode-collapse
on hard arithmetic. On \textbf{context rot} the DeepSeek pairs are
dominated by ceiling effects (instruct already at $1.000$ accuracy
across most cells, leaving no headroom), and the Qwen pairs show a
substantial reasoning advantage that is again partially entangled with
the same capability gap. We frame this as a \emph{selective}
intervention---reliably effective on hard single-shot inference,
heterogeneous on pushback resistance, and conferring no measurable
additional benefit on multi-turn state-tracking when the instruct
sibling is already at ceiling. The remainder of the paper
develops this in detail. Section~\ref{sec:related} positions the work
against the three axis-specific literatures; Section~\ref{sec:method}
explains the paired design and the four model pairs;
Section~\ref{sec:hypotheses} states the locked analysis plan;
Section~\ref{sec:setup} reports the exact sweep parameters and
exclusion accounting; Sections~\ref{sec:results}--\ref{sec:conclusion}
present results, discussion, and limitations.
