\section{Related Work}
\label{sec:related}

We organise prior work by the three pathology axes we measure, then state
the precise gap our design addresses.

\paragraph{Context rot.}
\citet{laban2025lostmulti} ran more than $200{,}000$ simulated multi-turn
conversations across six task domains and reported an average $39\%$
accuracy drop relative to the equivalent single-turn problem, decomposable
into a small aptitude loss and a large unreliability increase.
\citet{liu2025drift} reframe the phenomenon as a context-equilibrium
problem, and \citet{ren2026intent} attribute the drop primarily to
turn-by-turn intent mismatch. \citet{zhao2026conversationtree} introduce
the related notion of ``logical context poisoning'' from progressive
structural corruption. Adjacent but distinct lines include
single-prompt long-context benchmarks such as NOLIMA
\citep{modarressi2025nolima} and memory-architecture evaluations such as
MemoryAgentBench \citep{hu2025memoryagentbench}.

\paragraph{Sycophancy.}
\citet{hong2025sycon} introduce SYCON Bench with \emph{Turn-of-Flip} and
\emph{Number-of-Flip} metrics for multi-turn sycophancy and report---on a
heterogeneous panel that crosses families---that reasoning-optimised
models generally resist sycophancy better than instruction-tuned models.
\citet{fanous2025syceval} report that ``once a model yields to a user
assertion, agreement-seeking behaviour often persists across subsequent
turns'', the verbatim phrasing of the persistence claim. Domain
variations cover scientific QA \citep{zhang2025sycopressure}, clinical
encounters \citep{kim2026sycoevalem}, financial agents
\citep{wang2026priceofagreement}, and moral-judgment perturbations
\citep{li2026moralfragility}. \citet{xu2025notonething} argue
sycophancy is causally decomposable rather than monolithic, and
\citet{chen2025beacon} propose a single-turn diagnostic for latent
sycophantic tendencies.

\paragraph{Self-consistency under deterministic settings.}
\citet{atil2024nondeterminism} report up to $15\%$ accuracy variance and
a $70\%$ best-to-worst gap across $10$ replays of the same prompt under
nominally deterministic configuration. \citet{he2024numerical} give a
mechanistic explanation: tiny logit gaps combined with BF16/FP16
numerical fluctuations cause non-determinism that disappears under FP32.
\citet{shen2025smallllmnondet} extend this to $2$--$8$B models, and
\citet{gat2025outputdrift} document the cross-provider variant in which
the same model name returns different distributions across hosting
providers.

\paragraph{Agent-failure context.}
\citet{wang2025agentfail}, \citet{yu2025trajectbench}, and
\citet{lin2025agentracer} build failure taxonomies and localisation
methods for tool-using agents. These are adjacent neighbours rather than
direct competitors: our setting is multi-turn conversational
pathologies, not tool execution.

\paragraph{What is novel here.}
Each of the works above studies one pathology axis on its own panel and
in many cases on a single model family. Where reasoning-enabled variants
are implicated---most explicitly in SYCON \citep{hong2025sycon}---the
comparison crosses model families, so the reasoning-mode effect is
confounded with base architecture and scale. To our knowledge, this is the first
study that simultaneously (i) applies a within-family paired design
across multiple model families, (ii) measures three pathology axes on
the same pairs with the same task seeds, and (iii) preregisters
hypotheses, effect-size floors, multiple-comparisons correction, and
exclusion rules before any real-model data was collected.

\paragraph{Design-strength caveat by family.}
Within the within-family-paired framing, our four pairs are \emph{not}
controlled to the same degree, and we are explicit about which contrast
each pair supports. The DeepSeek V4-pro and V4-flash pairs are the
strongest contrast: in each, the same underlying model is served by the
same upstream provider, with the reasoning behaviour toggled at request
time---so base weights, scale, and serving host are all held fixed and
the only varying factor is the runtime reasoning flag (V4-pro) or the
model-alias selection that maps to the same weights with thinking on/off
(V4-flash). The Qwen3-235B and Qwen3-30B pairs are weaker: they match
base family and active-parameter count, but the instruct SKU and
thinking SKU are shipped as separate model identifiers and, in our
deployment, are served by \emph{different} upstream providers (Novita
vs Alibaba for Qwen3-235B; Nebius vs AtlasCloud for Qwen3-30B). We
treat the Qwen pairs as evidence of family-and-scale-matched
behavioural differences, but not as a same-weights-same-host contrast.
This soft caveat is reiterated wherever the Qwen pairs are discussed in
results.
