\section{Methodology}
\label{sec:method}

The design decision that organises everything else is the within-family
pairing. We describe the pairing first
(\S\ref{sec:method:pairing}), then the three tasks
(\S\ref{sec:method:tasks}), and finally the operational guarantees that
make the pairing meaningful---upstream-provider pinning, frozen
exclusion classes, resumable execution, and verifiable scoring
(\S\ref{sec:method:ops}).

\subsection{Within-family paired design}
\label{sec:method:pairing}

A canonical comparison such as ``does enabling a model's reasoning mode
change multi-turn behaviour?'' is usually answered by reading off
accuracy deltas across heterogeneous panels---e.g.\ a reasoning-enabled
model from lab A against an instruct model from lab B. Such a comparison
conflates the reasoning-mode contrast with base model, scale, data mix,
RLHF recipe, and serving stack. We avoid that confound by restricting attention to four
model \emph{families} that ship matched instruct and reasoning siblings:

\begin{enumerate}
  \item \textbf{DeepSeek V4-pro.} Same weights, two
        configurations of the reasoning toggle. The original sweep
        was pinned to Novita via OpenRouter
        (\texttt{reasoning: \{enabled: false\}} for instruct,
        \texttt{reasoning: \{effort: "high"\}} for reasoning); per the
        2026-05-15 amendment a subset of cells was recovered via the
        DeepSeek first-party API with the V4-family-native
        \texttt{thinking: \{type: "enabled"\|"disabled"\}} parameter.
        The final analyzable dataset therefore mixes OR/Novita
        originals with DS-direct recoveries on a per-cell basis; see
        \S\ref{sec:methods:v4pro-routing} for the rationale, accounting,
        and per-axis breakdown.
  \item \textbf{DeepSeek V4-flash.} Same weights, served by DeepSeek
        direct. The legacy aliases \texttt{deepseek-chat} and
        \texttt{deepseek-reasoner} route the same family through the
        instruct and reasoning code paths respectively.
  \item \textbf{Qwen3-235B (a22b).} Two SKUs published by Alibaba at the
        same active-parameter count: \texttt{qwen3-235b-a22b-2507}
        (instruct, served via Novita) and
        \texttt{qwen3-235b-a22b-thinking-2507} (reasoning, served via
        Alibaba).
  \item \textbf{Qwen3-30B (a3b).} \texttt{qwen3-30b-a3b-instruct-2507}
        (served via Nebius) and \texttt{qwen3-30b-a3b-thinking-2507}
        (served via AtlasCloud).
\end{enumerate}

For DeepSeek pairs base architecture, scale, training run, and serving
host are all held constant; only the reasoning toggle varies, so the
paired contrast isolates the reasoning effect. For Qwen pairs the
families and active-parameter counts match, but the default upstream
host differs across siblings (Novita vs.\ Alibaba on the 235B pair;
Nebius vs.\ AtlasCloud on the 30B pair). We document this as a soft
limitation of the Qwen pairing's serving-stack equivalence
(\S\ref{sec:limitations}) and report Qwen and DeepSeek results
separately so a reader can attribute any disagreement to the routing
difference.

This design is stronger than the pairing proposed in the original
preregistration (\texttt{deepseek-v4-pro} vs.\ \texttt{deepseek-r1-0528}),
which crossed two distinct base models. The amendment log dated
\textsl{2026-05-13 (further)} replaces that exploratory pairing with the
within-model reasoning-toggle design above, eliminating the
architecture/scale confound on the DeepSeek side.

\subsection{Three tasks, all integer-scored}
\label{sec:method:tasks}

All tasks emit a single integer answer that is extracted from the
model's response by a strict regular-expression scorer with
word-boundary matching. No LLM judge is invoked at any point in the
analysis loop.

\paragraph{Self-consistency: hard arithmetic (hardness $5$).}
The model is asked to evaluate a closed-form arithmetic expression in
$12$ four-digit operands with squares, modulo, integer division, and
nested grouping. Hardness $5$ was chosen so frontier models do not
saturate at $100\%$---which would leave no headroom for either
divergence or accuracy measurement. The same task seed is replayed
$N{=}25$ times at temperature $0$ with the API seed fixed at the
\texttt{task\_seed}, so any cross-replay divergence is attributable to
serving-stack non-determinism rather than nominal sampling randomness.
Per-task divergence is the fraction of replays whose extracted answer
differs from the modal extracted answer.

\paragraph{Context rot: variable tracking ($n_{\mathrm{updates}} = 20$).}
The model maintains a running integer state \texttt{x} across $20$
chained update turns---each turn applies an add, subtract, or multiply
operation---then is asked at probe for the final value of \texttt{x}.
This is a stronger context-rot signal than single-fact recall because
the probe is unanswerable without actually carrying state across turns.
Between the update block and the probe we insert $n_{\mathrm{filler}}
\in \{0, 2, 5, 10, 20, 40\}$ filler turn-pairs of one of four kinds:
\texttt{irrelevant} (chitchat), \texttt{related} (math-adjacent),
\texttt{token\_matched} (fixed-length filler at the same turn count
controlling for content variability), and \texttt{collapsed} (a single
fat user-assistant pair whose total token mass matches the cumulative
token mass of $n_{\mathrm{filler}}$ small pairs). The
\texttt{collapsed} kind is the true turn-vs-token control: comparing
accuracy at (\texttt{irrelevant}, $k{=}20$) against (\texttt{collapsed},
$k{=}20$) holds total tokens approximately constant while collapsing
$40$ turns into $2$, isolating turn-count effects from token-mass
effects.

\paragraph{Sycophancy: arithmetic-with-pushback (hardness $5$).}
The protocol is original probe $\rightarrow$ model answer $A$
$\rightarrow$ user pushback $\rightarrow$ model answer $A'$ $\rightarrow$
$g$ neutral filler pairs $\rightarrow$ re-probe $\rightarrow$ model
answer $A''$ (scored). The pushback condition is one of three: in the
\texttt{wrong} condition the user asserts a wrong integer answer (a
$+7$ perturbation of the truth that preserves magnitude and sign); in
the \texttt{correct} condition the user asserts the right answer; in
the \texttt{neutral} condition the user expresses doubt without
asserting any value. The \texttt{correct} condition is the critical
control: a model that flips on \texttt{correct} pushback is generically
capitulating, not sycophantic. The interpretable signal is
$\acc{\mathrm{correct}} - \acc{\mathrm{wrong}}$. The post-pushback gap
is $g \in \{0, 2, 5, 10\}$ neutral filler pairs.

The 2026-05-14 amendment log records the switch from a CRT-style
counterintuitive-math task to hardness-5 arithmetic. The reason was
empirical: stage-1 calibration showed that reasoning models solved CRT
items so reliably at baseline that wrong pushback produced essentially
no flip rate---a ceiling that hid any reasoning-vs-instruct difference.
Hardness-5 arithmetic creates legitimate uncertainty in model answers,
giving wrong pushback room to register.

\subsection{Operational guarantees}
\label{sec:method:ops}

\paragraph{Upstream-provider pinning.}
Every OpenRouter request specifies a single pinned upstream host
(\texttt{configs/models.yaml: upstream\_provider}) with
\texttt{allow\_fallbacks: false}. Without pinning, OpenRouter
load-balances across upstream hosts---which would silently confound a
paired test if one sibling was served by a different actual provider on
different days. After each completion we verify the response's reported
provider header against the pinned value and tag a mismatch as a
distinct exclusion class
(\texttt{upstream\_provider\_mismatch}).

\paragraph{Six frozen exclusion classes.}
Trajectories are excluded from analysis if and only if they fall into
one of six classes, all frozen alongside the preregistration:
(i) \texttt{provider\_error:*}: the request never reached the model
after retries (HTTP failure or credit exhaustion). Because the model
never produced output, these are re-attempted on resume rather than
counted as model-behaviour exclusions.
(ii) \texttt{refusal\_detected}: detected by absence of any digit in
the probe answer for arithmetic and variable-tracking tasks.
(iii) \texttt{truncated\_at\_max\_tokens}: missing terminating
punctuation \emph{and} output length equal to \texttt{max\_tokens}.
(iv) \texttt{upstream\_provider\_mismatch}: a fallback fired despite
\texttt{allow\_fallbacks: false}.
(v) \texttt{empty\_probe\_answer}: scorer found no integer in a
non-empty response (a true scoring failure).
(vi) \texttt{provider\_empty\_response}: the provider returned HTTP $200$
with an empty body and \texttt{output\_tokens == 0}, added in the
2026-05-14 amendment after stage-1 calibration revealed
${\sim}1{,}100$ such rows misclassified as \texttt{empty\_probe\_answer},
concentrated on \texttt{deepseek-reasoner} and \texttt{deepseek-v4-pro
reasoning} via Novita.

\emph{Originally we treated these as deterministic and not silently
re-attempted; the 2026-05-15 amendment retracts that judgment.}
Targeted diagnostics on 2026-05-15 showed the empties were the
downstream consequence of our \texttt{max\_tokens}{=}2048 setting being
below the DeepSeek-recommended default of $32{,}000$ for reasoning
models, so the entire budget was consumed by \texttt{reasoning\_content}
(CoT) and \texttt{content} came back empty. Affected cells were re-attempted
through DeepSeek's first-party API with
$\mathrm{max\_tokens}\in\{8{,}192, 16{,}384\}$ and recovered cleanly;
the analyzable dataset is composed of the recovered rows via semantic
dedup (\S\ref{sec:methods:dedup}). The original $2048$-token rows
survive in the JSONL audit trail. Classes (ii)--(vi) remain reported
on whatever residual fraction the recovery did not clear; only (i)
\texttt{provider\_error} continues to be silently re-attempted under
the runner's resume logic.

\paragraph{Resumable runner with content-hashed cell keys.}
Each trajectory is keyed by a content hash of (model, task seed,
sweep-cell coordinates). The JSONL log is appended in place. On resume
the runner skips any cell whose key already exists in the log. This
guarantees that a transient infrastructure failure mid-sweep cannot
silently corrupt a cell---the missing rows are simply re-attempted on
the next run, and only rows that produced a real model response are
counted.

\paragraph{Verifiable scoring, no LLM judges.}
All three tasks score a single extracted integer against a known
ground-truth integer using word-boundary regular expressions
(\texttt{src/agent\_pathologies/tasks/scoring.py}). This avoids the
known reliability issues with LLM-judge scoring on arithmetic-adjacent
tasks and makes the entire analysis pipeline reproducible from the raw
JSONL.

\paragraph{Paired by task seed.}
The paired statistical tests in Section~\ref{sec:hypotheses} are
constructed by matching trajectories on
$(\mathrm{family}, \mathrm{task\_seed}, \mathrm{sweep\_cell})$. The same
seed is run on both members of a pair, so the comparison conditions on
task identity and not just on marginal accuracy.
