\section{Experimental Setup}
\label{sec:setup}

This section reports the exact model identifiers, hyperparameters,
sweep dimensions, and exclusion accounting used to generate the
results.

\subsection{Models and upstream pinning}

Table~\ref{tab:models} lists the four within-family pairs with their
OpenRouter (or DeepSeek-direct) model identifiers, the pinned upstream
host, and the routing parameter that distinguishes the instruct member
from the reasoning member.

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{l l l l X}
\toprule
Family & Role & Model identifier & Upstream & Distinguishing parameter \\
\midrule
\multirow{2}{*}{\texttt{deepseek-v4-pro}$^\dagger$}
  & instruct  & \texttt{deepseek-v4-pro}\,/\,\texttt{deepseek/deepseek-v4-pro} & DS-direct\,/\,Novita & \texttt{thinking: \{type: "disabled"\}} (DS), \texttt{reasoning: \{enabled: false\}} (OR) \\
  & reasoning & \texttt{deepseek-v4-pro}\,/\,\texttt{deepseek/deepseek-v4-pro} & DS-direct\,/\,Novita & \texttt{thinking: \{type: "enabled"\}} (DS), \texttt{reasoning: \{effort: "high"\}} (OR) \\
\midrule
\multirow{2}{*}{\texttt{deepseek-v4-flash}}
  & instruct  & \texttt{deepseek-chat}     & DeepSeek direct & (instruct alias) \\
  & reasoning & \texttt{deepseek-reasoner} & DeepSeek direct & (reasoning alias) \\
\midrule
\multirow{2}{*}{\texttt{qwen3-235b}}
  & instruct  & \texttt{qwen/qwen3-235b-a22b-2507}          & Novita  & instruct SKU \\
  & reasoning & \texttt{qwen/qwen3-235b-a22b-thinking-2507} & Alibaba & thinking SKU \\
\midrule
\multirow{2}{*}{\texttt{qwen3-30b}}
  & instruct  & \texttt{qwen/qwen3-30b-a3b-instruct-2507} & Nebius     & instruct SKU \\
  & reasoning & \texttt{qwen/qwen3-30b-a3b-thinking-2507} & AtlasCloud & thinking SKU \\
\midrule
anchor (no pair) & instruct & \texttt{anthropic/claude-opus-4.7} & Anthropic & reported separately \\
\bottomrule
\end{tabularx}
\caption{Within-family pairs and upstream pinning. All OpenRouter
requests use \texttt{allow\_fallbacks: false}. Within each Qwen pair the
default upstream host differs across siblings (a soft serving-stack
limitation; see \S\ref{sec:limitations}). $^\dagger$ The
\texttt{deepseek-v4-pro} pair was originally pinned to OpenRouter via
Novita; per the 2026-05-15 PREREGISTRATION amendment, cells that
returned \texttt{provider\_error} or \texttt{empty\_probe\_answer} on
OR were re-attempted through DeepSeek's first-party API. The final
analyzable dataset is therefore \emph{mixed}: OR/Novita originals
that succeeded plus DS-direct recoveries for cells that the originals
did not produce. The slash-separated values list both
configurations; see \S\ref{sec:methods:v4pro-routing} for the per-axis
breakdown of the final mix.}
\label{tab:models}
\end{table}

\subsection{V4-pro routing amendment}
\label{sec:methods:v4pro-routing}

The \texttt{deepseek-v4-pro} pair was originally pinned to OpenRouter
via Novita, matching the upstream-pinning protocol applied to the other
OR-served families. Three independent observations on 2026-05-15
motivated a routing change to DeepSeek's first-party API
(\texttt{api.deepseek.com}) for all V4-pro retries:
(i)~\textbf{empty-response rate.}\ On long sycophancy and context-rot
conversations the OR-served V4-pro reasoning member returned a large
fraction of HTTP 200 responses with zero completion tokens, which our
runner correctly recorded as
\texttt{provider\_empty\_response} exclusions. Live testing against
\texttt{api.deepseek.com} with the documented native parameter
(\texttt{thinking: \{type: "enabled"\}}) and a budget of $16{,}384$
tokens reproducibly returned a populated \texttt{content} field on the
same prompts. (ii)~\textbf{the V4-family-native parameter is not
honoured for the legacy \texttt{reasoning} keyword.}\ The DeepSeek
documentation states explicitly that the V4 family uses
\texttt{thinking: \{type: "enabled"\|"disabled"\}}; the legacy
\texttt{reasoning: \{enabled: bool\}} is silently ignored. Empirically
we verified this on a controlled prompt: V4-pro responses with
\texttt{reasoning: \{enabled: false\}} continued to emit non-zero
reasoning tokens, whereas \texttt{thinking: \{type: "disabled"\}}
produced reasoning tokens of zero. (iii)~\textbf{methodological
parity}: V4-flash already used DeepSeek-direct, so routing V4-pro the
same way places both V4-family pairs on the model creator's API and
removes a confound between the two pairs. The 75\% promotional
discount on V4-pro through 2026-05-31 made the routing change financially
no worse than continuing on OR.

The retry rows written through DeepSeek-direct supersede the
\texttt{provider\_error} / \texttt{provider\_empty\_response} originals
via the semantic dedup at analysis time
(\S\ref{sec:methods:dedup}). Both rows remain in the raw JSONL
audit trail; the analyzable set picked by
\texttt{dedupe\_to\_latest} is composed of OR/Novita originals where
they succeeded, plus DS-direct retries on cells where the OR-served
original failed. Concretely (counts after dedup): on self-consistency,
$820$ of $1{,}000$ V4-pro-instruct and $400$ of $1{,}000$ V4-pro-reasoning
cells are OR-original; $180$ and $600$ respectively are DS-direct
recoveries. On sycophancy, $600$ of $600$ V4-pro-instruct cells are
OR-original; $260$ of $600$ V4-pro-reasoning are OR-original and the
other $340$ are DS-direct recoveries. On context-rot, the mix is
strongly OR-original: $1{,}187$ of $1{,}200$ V4-pro-instruct and
$1{,}168$ of $1{,}200$ V4-pro-reasoning come from the original Novita
sweep; the small DS-direct contributions ($13$ and $32$) are
recoveries on the cells where Novita returned errors or empty bodies.
The mix is heaviest on the reasoning sycophancy cells precisely
because that was the cell type most affected by the
\texttt{max\_tokens}{=}2048 bug.

\subsection{Hyperparameters per experiment}

Table~\ref{tab:hparams} reports temperature, max-tokens, sweep
dimensions, and total trajectories per family for each of the three
experiments. Temperature is fixed at $0$ on every call; the API seed is
held at the task seed for the self-consistency replays so that
remaining variance is attributable to serving-stack non-determinism.
The initial sweep used a \texttt{max\_tokens} budget of $2048$, which
turned out to be below the DeepSeek-recommended default of $32{,}000$ for
the \texttt{deepseek-reasoner} alias and caused systematic
empty-content responses on reasoning calls whose CoT exhausted the
budget. Per the 2026-05-15 PREREGISTRATION amendment, all retries on
DeepSeek reasoning cells used $\mathrm{max\_tokens}{=}8{,}192$ for
self-consistency / sycophancy and $\mathrm{max\_tokens}{=}16{,}384$ for
context-rot (whose longer input conversations leave less of the
\texttt{completion\_tokens} budget for content after reasoning); Qwen
instruct retries used $8{,}192$ to fix a smaller truncation issue on
$h{=}5$ arithmetic. The final analyzable dataset is composed of the
highest-token successful attempt per logical cell via the semantic
dedup (\S\ref{sec:methods:dedup}). The original $2048$ rows survive in
the JSONL audit trail.

\begin{table}[h]
\centering
\small
\begin{tabular}{l l l l l}
\toprule
Experiment & Task & Sweep dimensions & Per-pair $N$ & Hyperparameters \\
\midrule
self-consistency
  & arithmetic ($h{=}5$)
  & $40$ tasks $\times 25$ replays
  & $2{,}000$
  & $T{=}0$, max tok $2048$ \\
context rot
  & variable\_tracking ($n{=}20$)
  & $50 \times 6 \times 4$
  & $2{,}400$
  & $T{=}0$, max tok $2048$ \\
sycophancy
  & arithmetic ($h{=}5$)
  & $50 \times 3 \times 4$
  & $1{,}200$
  & $T{=}0$, max tok $2048$ \\
\bottomrule
\end{tabular}
\caption{Hyperparameters and sweep dimensions per experiment.
Per-pair $N$ counts trajectories collected on \emph{both} members of
the pair, so an effective paired sample size is half of this when
matched by task seed.}
\label{tab:hparams}
\end{table}

\subsection{Exclusion-rule definitions}

The six exclusion classes (\S\ref{sec:method:ops}) are defined formally
in \texttt{src/agent\_pathologies/analysis/exclusions.py} and applied at
analyzer read-time. Class~(i), \texttt{provider\_error}, is
infrastructure-only and re-attempted on resume; classes~(ii)--(vi) are
model-behaviour or serving-stack exclusions and are reported as-is.
Legacy rows tagged \texttt{empty\_probe\_answer} prior to the
2026-05-14 amendment are retroactively reclassified by
\texttt{metrics.py::reclassify\_legacy\_exclusions} based on the
\texttt{output\_tokens} field, idempotently.

\subsection{Semantic deduplication}
\label{sec:methods:dedup}

The raw JSONL is append-only: every retry attempt writes a new row,
even when retrying the same logical (model-family, role, task,
sweep-cell, seed) cell. For the analyzable dataset we apply
\texttt{dedupe\_to\_latest} (\texttt{src/agent\_pathologies/analysis/metrics.py}),
which keys cells on \emph{(model-family, role, task, sweep, seed)} —
intentionally broader than the runner's resume-key, which additionally
distinguishes by the raw model identifier. This broader key collapses
rows of the same logical cell that were served by different providers
(notably V4-pro originally via OR/Novita and re-served via
DeepSeek-direct in the retries), preferring any non-excluded row over
any excluded row, and tie-breaking on the latest non-excluded row. The
result is exactly one row per logical cell in the analyzed dataset
($N_{\mathrm{analyzable}}$ in Table~\ref{tab:pairs}). All raw rows
remain in the JSONL for audit.

\subsection{Statistical pipeline}

The analysis pipeline is fully deterministic given the JSONL trajectory
logs. Bootstrap CIs use $10{,}000$ resamples with a seeded random
state. McNemar exact tests use the exact binomial form to avoid the
continuity-correction debate at small discordance counts. BH correction
is applied across all cells within each axis, separately for each
co-primary metric in the self-consistency and sycophancy experiments.

\subsection{Reproducibility}

All raw JSONL trajectories will be released alongside the paper, along
with the analysis CSVs (\texttt{data/*\_paired.csv},
\texttt{pair\_analysis.csv}) that the figures and tables are produced
from. Re-running \texttt{scripts/run\_sweep.sh} reproduces the data
from the same configuration files; re-running
\texttt{experiments/<axis>/analyze.py} reproduces the figures from the
data.
