\section{Limitations}
\label{sec:limitations}

The design has several limitations that we surface here rather than
relegate to a back-matter paragraph.

\paragraph{Capability/sycophancy confound on Qwen.}
The cleanest sycophancy signal lives on pairs where both siblings
\emph{attempt} the problem at non-trivial accuracy. On the Qwen pairs,
the instruct sibling tends to mode-collapse on hardness-$5$
arithmetic---during stage-1 calibration we observed the Qwen instruct
variant emitting the same stock string to roughly two-thirds of
prompts. When the instruct baseline accuracy is near zero, the
reasoning-versus-instruct contrast on $\acc{\mathrm{correct}} -
\acc{\mathrm{wrong}}$ entangles two effects: a genuine sycophancy
difference and a capability gap that prevents the instruct member from
ever being in a position to flip. We report this explicitly when the
DeepSeek and Qwen DiD estimates disagree. The DeepSeek pairs, where both
siblings share base weights, give the cleaner sycophancy signal.

\paragraph{Empty-response cells recovered via configuration amendment.}
The original sweep produced a substantial
\texttt{empty\_probe\_answer} / \texttt{provider\_empty\_response}
class concentrated on DeepSeek reasoning members
($\approx\!668$ on \texttt{deepseek-reasoner} and $\approx\!407$ on
\texttt{deepseek-v4-pro} reasoning during stage-1 calibration), at first
treated as a model-side behavioural signal. Diagnostic work on
2026-05-15 traced these rows to an experimenter-configuration error:
\texttt{max\_tokens}{=}2048 was below DeepSeek's recommended default of
32K for reasoning models, so the entire budget was consumed by
\texttt{reasoning\_content} (CoT), leaving zero tokens for
\texttt{content} (the final visible answer). Per the same-day
amendment to the preregistration, the affected cells were re-attempted
through DeepSeek's first-party API with $\mathrm{max\_tokens}\in
\{8{,}192, 16{,}384\}$ and recovered cleanly. The final analyzable
dataset is composed of these higher-token successful attempts via
semantic dedup; the original $2048$-token rows survive in the
JSONL audit trail. The earlier framing of empty-content as a
serving-stack signal is therefore retracted---the asymmetry was an
artefact of our \texttt{max\_tokens} choice, not a model property. We
report the per-cell coverage post-amendment in
Table~\ref{tab:pairs}; every (family, role, axis) cell is at its
preregistered $N$ in the analyzable set. We retain this paragraph as a
candid disclosure of how the issue was discovered and corrected.

\paragraph{Single difficulty level on arithmetic axes.}
Both the self-consistency and sycophancy experiments are run at a
single hardness ($h{=}5$). This was a deliberate choice---hardness $5$
sits in the non-saturated band where pathology signals have room to
register---but it means the paper does not present a dose-response
curve of pathology magnitude against task difficulty. Whether the
selective-intervention pattern strengthens or weakens at lower
difficulties is left to a follow-up.

\paragraph{No closed-frontier anchor in the primary results.}
A stage-2 anchor on \texttt{anthropic/claude-opus-4.7} is provisioned
in the configuration but is not part of the paired test
(\S\ref{sec:method:pairing}). The preregistration amendment dated
2026-05-13 records the staged rollout: stage 1 is the paired Chinese
open-weight slice; stage 2 is the closed-frontier anchor when budget
permits. The headline within-family claim is independent of the anchor.

\paragraph{Within-Qwen-pair upstream-provider difference.}
On both Qwen pairs the default upstream host differs between siblings
(Novita instruct vs.\ Alibaba thinking on $235$B; Nebius instruct vs.\
AtlasCloud thinking on $30$B). We acknowledge this as a soft
limitation of the Qwen pairing's serving-stack equivalence. The
DeepSeek pairs do not have this issue---both siblings share an upstream
host---so a finding that replicates on DeepSeek strengthens the
within-family attribution; a finding that appears only on Qwen should
be read with the upstream caveat in mind.

\paragraph{Behavioural-only.}
The paper measures input-output behaviour. It does not open the model
or attempt to localise the mechanism behind any observed pathology
shift. Mechanistic follow-up on an open-weight pair (e.g.\ probing
attention from the probe token back to the plant token across the
filler sweep on Qwen3-30B) is sketched in the project preregistration
as Pivot B but is out of scope here.

\paragraph{Domain coverage.}
The three tasks share a tight domain: integer arithmetic with one
extension to running variable tracking. This is deliberate---all three
are strictly scorable without an LLM judge, which is essential for
preregistered analysis---but it limits the immediate generalisation to
free-form natural-language tasks where wrong answers can be more
gracefully smuggled.

\paragraph{Provider scope.}
Within-family pairs are restricted to Chinese open-weight families
because, at time of writing, those are the families that ship matched
instruct and reasoning siblings at the same scale. We do not claim the
within-family pattern generalises to closed-frontier reasoning models
without further data.
