\section{Reproducibility}
\label{sec:reproducibility}

We commit to reproducibility at the per-trajectory level. The companion
repository ships the locked preregistration, all amendments with dates,
the exact configuration files, the data-collection harness, the
statistical analysis pipeline, and all raw and cleaned trajectory logs
under permissive licensing.

\paragraph{Code and data release.}
\begin{itemize}
  \item \textbf{Repository:}
        \texttt{github.com/t4n15hq/agent-pathologies}
  \item \textbf{License:} CC BY 4.0 for the paper and accompanying
        data; MIT for the code
  \item \textbf{Data:} the three \texttt{data/*\_clean.jsonl} files (one
        per experiment) carry the analyzable dataset after semantic
        deduplication (\S\ref{sec:methods:dedup}); the raw
        \texttt{data/*.jsonl} files carry the full audit trail
        including superseded retry rows; the \texttt{figures/} PDFs
        are in the release archive.
\end{itemize}

\paragraph{Configuration and amendments.}
\begin{itemize}
  \item \texttt{configs/models.yaml} (within-family pairs, upstream pins)
  \item \texttt{configs/pivot\_a.yaml} (sweep parameters: $n_{\mathrm{tasks}}$,
        replays, filler counts/kinds, conditions/gaps, $T = 0$,
        \texttt{max\_tokens} per axis)
  \item \texttt{PREREGISTRATION.md} (root) and
        \texttt{experiments/*/PREREGISTRATION.md} (per-experiment) with
        the dated amendment log. Every amendment in the log is reflected
        in the final code state.
\end{itemize}

\paragraph{Run metadata.}
\begin{itemize}
  \item \textbf{Data-collection dates:} 2026-05-13 to 2026-05-15,
        including the initial sweep and the three amendments (cell-key
        fix on 2026-05-14; \texttt{max\_tokens} bug-fix and V4-pro
        routing change on 2026-05-15).
  \item \textbf{Trajectory counts (raw / analyzable / superseded-by-retry):}
        self-consistency $10{,}573$ / $8{,}000$ / $2{,}573$;
        sycophancy $5{,}467$ / $4{,}800$ / $667$;
        context rot $15{,}862$ / $9{,}600$ / $6{,}262$. Every
        (family, role, axis) cell is at its preregistered $N$ in the
        analyzable set.
  \item \textbf{Total cost:} approximately \$170 USD (OpenRouter spend
        \$155, DeepSeek-direct \$15) across the four Chinese
        open-weight families.
\end{itemize}

\paragraph{Regenerating the analyses.}
From the repository root:
\begin{verbatim}
# 1. Set up the environment
python -m venv .venv && source .venv/bin/activate
pip install -e ".[dev]"

# 2. (Optional) Run the post-sweep clean-up passes
python scripts/retry_truncated.py --apply
python scripts/clean_dataset.py

# 3. Per-axis paired tests + cell-level summaries
python experiments/self_consistency/analyze.py
python experiments/context_rot/analyze.py
python experiments/sycophancy/analyze.py

# 4. Cross-axis headline used in the paper
python scripts/pair_analysis.py
\end{verbatim}
Each analyser is deterministic given the input JSONL; bootstrap CIs and
permutation tests use a fixed seed (\texttt{seed=42}) hard-coded in
\texttt{src/agent\_pathologies/analysis/stats.py}.

\paragraph{File hashes.}
For audit purposes, the SHA-256 hashes of the released cleaned JSONL
files are:
\begin{itemize}
  \item \texttt{data/self\_consistency\_clean.jsonl}: \texttt{ff9eecc6\dots f5f0ddb3}
  \item \texttt{data/context\_rot\_clean.jsonl}: \texttt{61c1a4fb\dots 0c0852751}
  \item \texttt{data/sycophancy\_clean.jsonl}: \texttt{5f36a88f\dots 0108371b}
\end{itemize}
Full hashes are pinned in the release tag and reproducible with
\texttt{shasum -a 256 data/*\_clean.jsonl} after running
\texttt{python scripts/clean\_dataset.py}.

\paragraph{Provider-pinning verification.}
Every trajectory in the released JSONL carries an
\texttt{extra.upstream\_pinned} field and an
\texttt{extra.upstream\_actual} field. Exclusion rule \S6.4
(\texttt{upstream\_mismatch}) drops any trajectory where the two differ.
A reader can audit the entire dataset for routing correctness with one
\texttt{jq} query against the JSONL.

