Multi-turn pathologies in large language models---self-consistency drift,
context rot, and sycophancy---have each been documented in isolation, but
the field lacks a controlled, preregistered comparison of how
\emph{reasoning-enabled variants} of a model differ from instruct variants
within a fixed family. We address this gap with a paired, multi-axis
study of four Chinese open-weight model families---DeepSeek V4-pro,
DeepSeek V4-flash, Qwen3-235B, and Qwen3-30B---in which each family
contributes an instruct mode and a reasoning mode. The DeepSeek pairs
isolate the cleanest possible contrast: the \emph{same} underlying model
served by the \emph{same} upstream, toggled between reasoning on and
reasoning off at request time. The Qwen pairs match base family and active
parameter count, but the instruct and thinking SKUs are served by different
upstream providers, which we treat as a soft caveat throughout. Hypotheses,
primary metrics, effect-size thresholds, multiple-comparisons correction,
exclusion rules, and falsification criteria were locked before any
real-model trajectory was collected. We measured behaviour on three
verifiable, integer-scored tasks (hard arithmetic, $20$-update variable
tracking, and arithmetic-with-pushback) under matched paired sweeps,
upstream-provider pinning, and frozen exclusion classes. Reasoning-enabled
variants show large and universal self-consistency advantages
($\Delta_{\mathrm{acc}}$ of $+0.67$ to $+0.99$ across the four pairs,
sign-test BH-$q < 10^{-11}$ in every family); on sycophancy the
pattern is heterogeneous within the family of reasoning-enabled
variants, with a clean within-MODEL effect on V4-flash
(reasoning-gain $+0.165$, BH-$q \approx 5{\times}10^{-4}$), a null
at threshold on V4-pro (gain $+0.045$, CI $[-0.06, +0.15]$), and
larger but capability-confounded gains on the Qwen pairs. Context rot is
dominated by ceiling effects on the DeepSeek pairs and shows a
substantial reasoning advantage on Qwen that is partially entangled
with the same instruct-side mode-collapse. We interpret reasoning
enablement as a \emph{selective} intervention---reliably effective on
hard single-shot inference, heterogeneous on pushback resistance, and
conferring no measurable additional benefit on multi-turn
state-tracking when the instruct sibling is already at ceiling.
